Structured Versus Decorated Cospans

### 2020-11-10 cospans.tex
typo fixed
decommented introduction

\section{Introduction}
\begin{comment}
Networks are playing an increasingly prominent role in our understanding of the world and as a result, methods of characterizing and studying networks have become evermore necessary. Applied category theory provides such an avenue and much work has been done towards this effort  \cite{BC,BCR,BF,BFP,BP,Cour,Fong,JM,LS,Yass}. One of the more recent frameworks developed that is suitable for studying networks is Fong's `decorated cospans' \cite{Fong}. A $\textbf{cospan}$ in any category is a diagram of the form:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a_1$};
\node (B) at (1,1) {$b$};
\node (C) at (2,0) {$a_2$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
We call $b$ the $\textbf{apex}$ of the cospan and $a_1$ and $a_2$ the $\textbf{feet}$ of the cospan. The morphisms $i$ and $o$ are called the $\textbf{legs}$. Cospans are ideal for realizing networks and more generally `open' systems where here open means that each system or network comes with prescribed inputs and outputs which are represented by the feet of the cospan $a$ and $b$, respectively. Two open systems viewed as cospans such that the outputs of the first and the inputs of the second coincide can then naturally be composed via pushout to obtain an open system whose inputs are that of the first and outputs are that of the second. For example, we can compose the above cospan with the following:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a_2$};
\node (B) at (1,1) {$b^\prime$};
\node (C) at (2,0) {$a_3$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i^\prime$} (B)
(C) edge node[above]{$o^\prime$} (B);
\end{tikzpicture}
\]
to obtain a new cospan whose left foot is the left foot of the first, whose right foot is the right foot of the second, and whose apex and legs are given by the pushout.
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a_1$};
\node (B) at (1,1) {$b$};
\node (C) at (2,0) {$a_2$};
\node (D) at (3,1) {$b^\prime$};
\node (E) at (4,0) {$a_3$};
\node (F) at (2,2) {$b+_{a_2} b^\prime$};
\node (G) at (2,1) {$b+b^\prime$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(C) edge node[above]{$i^\prime$} (D)
(E) edge node[above]{$o^\prime$} (D)
(B) edge node[left]{$\psi j_b$} (F)
(D) edge node[right]{$\psi j_{b^\prime}$} (F)
(B) edge node[above]{$j_b$} (G)
(D) edge node[above]{$j_{b^\prime}$} (G)
(G) edge node [left] {$\psi$} (F);
\end{tikzpicture}
\]

@@

\subsection*{Acknowledgements}
Daniel Cicala, various others.
\end{comment}

@@

\begin{cor}
Let $\A$ be a category with finite colimits and $F \maps (\A, +) \to \Cat$ a symmetric lax monoidal pseudofunctor.  Thenthere exists a symmetric monoidal category $F\textnormal{Csp}$ which has:
\begin{enumerate}
\item{objects as those of $\mathsf{A}$ and}
\item{morphisms as isomorphism classes of $F$-decorated cospans of $\mathsf{A}$, where an $F$-decorated cospan is given by a pair:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$c$};
\node (C) at (2,0) {$b$};
\node (D) at (4,0) {$d \in F(c)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
Given another $F$-decorated cospan:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$c^\prime$};
\node (C) at (2,0) {$b$};
\node (D) at (4,0) {$d^\prime \in F(c^\prime)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[below]{$i^\prime$} (B)
(C) edge node[below]{$o^\prime$} (B);
\end{tikzpicture}
\]
these two $F$-decorated cospans are in the same isomorphism class if there exists an isomorphism $f \maps c \to c^\prime$ such that following diagram commutes:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B') at (1,0.5) {$c$};
\node (B) at (1,-0.5) {$c^\prime$};
\node (C) at (2,0) {$b$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[below]{$i^\prime$} (B)
(C) edge node[below]{$o^\prime$} (B)
(A) edge node[above]{$i$} (B')
(C) edge node[above]{$o$} (B')
(B') edge node[left]{$f$} (B);
\end{tikzpicture}
\]
and there exists an isomorphism $\iota \maps F(f)(d) \to d^\prime$ in $F(c^\prime)$.}
\end{enumerate}
\end{cor}
In this symmetric monoidal category, isomorphism classes are as they should morally be, and the instance of two graphs having different edge sets does not prevent them from being in the same isomorphism class due to the isomorphism $\iota$.

>>>>

\section{Introduction}
Networks are playing an increasingly prominent role in our understanding of the world and as a result, methods of characterizing and studying networks have become evermore necessary. Applied category theory provides such an avenue and much work has been done towards this effort  \cite{BC,BCR,BF,BFP,BP,Cour,Fong,JM,LS,Yass}. One of the more recent frameworks developed that is suitable for studying networks is Fong's `decorated cospans' \cite{Fong}. A $\textbf{cospan}$ in any category is a diagram of the form:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a_1$};
\node (B) at (1,1) {$b$};
\node (C) at (2,0) {$a_2$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
We call $b$ the $\textbf{apex}$ of the cospan and $a_1$ and $a_2$ the $\textbf{feet}$ of the cospan. The morphisms $i$ and $o$ are called the $\textbf{legs}$. Cospans are ideal for realizing networks and more generally `open' systems where here open means that each system or network comes with prescribed inputs and outputs which are represented by the feet of the cospan $a$ and $b$, respectively. Two open systems viewed as cospans such that the outputs of the first and the inputs of the second coincide can then naturally be composed via pushout to obtain an open system whose inputs are that of the first and outputs are that of the second. For example, we can compose the above cospan with the following:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a_2$};
\node (B) at (1,1) {$b^\prime$};
\node (C) at (2,0) {$a_3$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i^\prime$} (B)
(C) edge node[above]{$o^\prime$} (B);
\end{tikzpicture}
\]
to obtain a new cospan whose left foot is the left foot of the first, whose right foot is the right foot of the second, and whose apex and legs are given by the pushout.
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a_1$};
\node (B) at (1,1) {$b$};
\node (C) at (2,0) {$a_2$};
\node (D) at (3,1) {$b^\prime$};
\node (E) at (4,0) {$a_3$};
\node (F) at (2,2) {$b+_{a_2} b^\prime$};
\node (G) at (2,1) {$b+b^\prime$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(C) edge node[above]{$i^\prime$} (D)
(E) edge node[above]{$o^\prime$} (D)
(B) edge node[left]{$\psi j_b$} (F)
(D) edge node[right]{$\psi j_{b^\prime}$} (F)
(B) edge node[above]{$j_b$} (G)
(D) edge node[above]{$j_{b^\prime}$} (G)
(G) edge node [left] {$\psi$} (F);
\end{tikzpicture}
\]

@@

\subsection*{Acknowledgements}
Daniel Cicala, various others.

@@

\begin{cor}
Let $\A$ be a category with finite colimits and $F \maps (\A, +) \to \Cat$ a symmetric lax monoidal pseudofunctor.  Then there exists a symmetric monoidal category $F\textnormal{Csp}$ which has:
\begin{enumerate}
\item{objects as those of $\mathsf{A}$ and}
\item{morphisms as isomorphism classes of $F$-decorated cospans of $\mathsf{A}$, where an $F$-decorated cospan is given by a pair:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$c$};
\node (C) at (2,0) {$b$};
\node (D) at (4,0) {$d \in F(c)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
Given another $F$-decorated cospan:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$c^\prime$};
\node (C) at (2,0) {$b$};
\node (D) at (4,0) {$d^\prime \in F(c^\prime)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[below]{$i^\prime$} (B)
(C) edge node[below]{$o^\prime$} (B);
\end{tikzpicture}
\]
these two $F$-decorated cospans are in the same isomorphism class if there exists an isomorphism $f \maps c \to c^\prime$ such that following diagram commutes:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B') at (1,0.5) {$c$};
\node (B) at (1,-0.5) {$c^\prime$};
\node (C) at (2,0) {$b$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[below]{$i^\prime$} (B)
(C) edge node[below]{$o^\prime$} (B)
(A) edge node[above]{$i$} (B')
(C) edge node[above]{$o$} (B')
(B') edge node[left]{$f$} (B);
\end{tikzpicture}
\]
and there exists an isomorphism $\iota \maps F(f)(d) \to d^\prime$ in $F(c^\prime)$.}
\end{enumerate}
\end{cor}
In this symmetric monoidal category, isomorphism classes are as they should morally be, and the instance of two graphs having different edge sets does not prevent them from being in the same isomorphism class due to the isomorphism $\iota$.
